---
title: "Data Analysis Religiosity's Part on Social Adaptation"
author: "Alejandra Neely-Prado, Michiel van Elk, Gorka Navarrete, Fernanda Hola, David Huepe"
date: "12-09-2021"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
params:
  depends: ''
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

# Preparation old data 

DF_old = 
  read_csv(here::here("data/2017_experimental_group.csv"), 
           col_types = 
             cols(
              .default = col_double(),
              ISD07 = col_character(),
              ISD04 = col_character(),
              ISD05 = col_character(),
              ISD14 = col_character(),
              ISD28 = col_character(),
              EAR03 = col_character(),
              EAR05 = col_character(),
              EAR08 = col_character(),
              EAR09 = col_character(),
              EAR10 = col_character())) %>% 
            filter(
    grepl("chile|chileno|chilena|Chile", ISD04, ignore.case = TRUE) &
      ISD08 >= 18 &
      as.numeric(stringr::str_extract(ISD05, "[0-9]{1,3}")) >= 10) %>% 
  drop_na(ISD07, ISD08, Religiosidad, Autoestima, SASS_Total) %>% 
  select(ISD07, ISD04, ISD05, ISD08, ISD14, ISD28,
         Reli01:Reli10, Religiosidad,
         EAR01, EAR02, EAR03, EAR04, EAR05,EAR06,EAR07,EAR08,EAR09,EAR10, Autoestima,
         SASS01:SASS21, SASS_Total, Estres_Percibido) %>% 
  filter(grepl("Chile|Chilena|chile|chilena|chileno|Chileno", ISD04))

DF_old[DF_old=="Muy en desacuerdo"]<-"4"
DF_old[DF_old=="En desacuerdo"]<-"3"
DF_old[DF_old=="De acuerdo"]<-"2"
DF_old[DF_old=="Muy de acuerdo"]<-"1"

proy_no_na_old = DF_old %>% 
  rowwise() %>% 
  mutate(
    anios_chile = as.integer(substr(ISD05, 1, 2)),
    age = ISD08,
    sex = ifelse(ISD07 == "Masculino", 1, 0),
    edu_level = as.numeric(gsub("([0-9]+).*$", "\\1", ISD14))) %>%
  mutate(
    EAR03 = as.numeric(EAR03),
    EAR05 = as.numeric(EAR05),
    EAR08 = as.numeric(EAR08),
    EAR09 = as.numeric(EAR09),
    EAR10 = as.numeric(EAR10),
    SASS00 = NA,
  ) %>% 
  filter(age >= 18 & anios_chile >= 10) %>% 
  select(-contains("ISD"), -anios_chile) %>% 
  mutate(group = "EX") 

```

```{r}
# Preparation new data ----------------------------------------------------
id_match = read_csv(here::here("outputs/data/DF_matched.csv"), 
                    col_types = 
                      cols(
                        id_old = col_double(),
                        DEMOGR3_01_DIR = col_double(),
                        DEMOGR3_02_DIR = col_double(),
                        ROW = col_double(),
                        id = col_character())) %>% pull(id)

data_new = read_csv(here::here("outputs/data/DF_joined.csv"), col_types = cols()) %>% filter(id %in% id_match)

DF_new = data_new %>% transmute(
  id = id,
  sex = DEMOGR3_02_DIR, # 1 = Hombre; 0 = Mujer ; 2 = otro
  age = DEMOGR3_01_DIR,
  edu_level = DEMOGR3_05_DIR,
  criterio_exclusion_nac =  DEMOGR3_03_DIR, # ¿Usted nació en Chile?
  criterio_exclusion_nac10 =  DEMOGR3_04_DIR, # ¿Usted vive hace más de 10 años en Chile?
  criterio_exclusion_psi = DEMOGR3_07_DIR,
  criterio_exclusion_fis = DEMOGR3_08_DIR,
  clasificacion_gse = AIM_DIRt,
  Reli01 = SCSORF_01_DIR,
  Reli02 = SCSORF_02_DIR,
  Reli03 = SCSORF_03_DIR,
  Reli04 = SCSORF_04_DIR,
  Reli05 = SCSORF_05_DIR,
  Reli06 = SCSORF_06_DIR,
  Reli07 = SCSORF_07_DIR,
  Reli08 = SCSORF_08_DIR,
  Reli09 = SCSORF_09_DIR,
  Reli10 = SCSORF_10_DIR,
  EAR01 = EAR_01_DIR,
  EAR02 = EAR_02_DIR,
  EAR03 = EAR_03_DIR,
  EAR04 = EAR_04_DIR,
  EAR05 = EAR_05_DIR,
  EAR06 = EAR_06_DIR,
  EAR07 = EAR_07_DIR,
  EAR08 = EAR_08_DIR,
  EAR09 = EAR_09_DIR,
  EAR10 = EAR_10_DIR,
  SASS00 = SASS_01_DIR,
  SASS01 = SASS_02_DIR,
  SASS02 = SASS_03_DIR,
  SASS03 = SASS_04_DIR,
  SASS04 = SASS_05_DIR,
  SASS05 = SASS_06_DIR,
  SASS06 = SASS_07_DIR,
  SASS07 = SASS_08_DIR,
  SASS08 = SASS_09_DIR,
  SASS09 = SASS_10_DIR,
  SASS10 = SASS_11_DIR,
  SASS11 = SASS_12_DIR,
  SASS12 = SASS_13_DIR,
  SASS13 = SASS_14_DIR,
  SASS14 = SASS_15_DIR,
  SASS15 = SASS_16_DIR,
  SASS16 = SASS_17_DIR,
  SASS17 = SASS_18_DIR,
  SASS18 = SASS_19_DIR,
  SASS19 = SASS_20_DIR,
  SASS20 = SASS_21_DIR,
  SASS21 = SASS_22_DIR,
  Autoestima = EAR_DIRt,
  SASS_Total = SASS_DIRt,
  Religiosidad = SCSORF_DIRt,
  Estres_Percibido = PSS_DIRt) %>% 
  distinct(id, .keep_all = TRUE) %>% 
  select(-contains("criterio"), -"clasificacion_gse", -"id") %>% mutate(group = "CN")

matched = DF_new %>% rbind(proy_no_na_old)%>% rename(
  "Religiosity" = "Religiosidad",
  "Self-Esteem" = "Autoestima",
  "Social-Adaptation" = "SASS_Total",
  "Age" = "age")

```


This document contains the pre-registered analysis plan, and also an exploratory analysis.

# Pre-registered analysis plan

## Statistical Analysis

### Reliability

To start the analysis, we will get Cronbach's alpha to check each questionnaires reliability. In general all questionnaires show good levels of reliability.

#### Religiosity 

```{r}
rel = matched %>% select(Reli01:Reli10) 
psych::alpha(rel, check.keys = TRUE)[1] %>% kable()
```

#### Social Adaptation

```{r}
sass = matched %>% mutate(
  sass_01final = ifelse(!is.na(SASS01), SASS01, SASS02)
) %>% select(sass_01final, SASS03:SASS21)
psych::alpha(sass, check.keys = TRUE, warnings = FALSE)[1] %>% kable()
```

#### Self-Esteem

```{r}
aut = matched %>% select(EAR01:EAR10)
psych::alpha(aut)[1] %>% kable
```

## Results

### Descriptives 

```{r}
sex_table = table(matched$sex)
females_n = sex_table[1]
males_n = sex_table[2]
total = females_n + males_n

age_mean_females = matched %>% filter(sex == 0) %>% select(Age) %>% pull() %>% mean()
age_mean_males = matched %>% filter(sex == 1) %>% select(Age) %>% pull() %>% mean()

```

`r females_n` females (`r round(females_n/total*100, 2)`%) and `r males_n` males (`r round(males_n/total*100, 2)`%) enrolled in this study, with an average age of `r round(age_mean_females, 2)` and `r round(age_mean_males, 2)`, respectively. 

The following table show the descriptive statistics for each independent and dependent variable.

```{r}
### TABLE 1(mean, sd, variance, range, assymetry, kurtosis). This should be the table we insert to the paper
for_descriptives = matched %>% select(Religiosity, `Social-Adaptation`, `Self-Esteem`, group)
descriptives = psych::describeBy(for_descriptives %>% select(Religiosity, `Social-Adaptation`, `Self-Esteem`), group = for_descriptives$group)
descriptives
```

### Correlations


```{r include=FALSE}

CORREL = correlation::correlation(matched %>% select(Religiosity, `Social-Adaptation`, `Self-Esteem`))

CORREL_RE_SO = CORREL %>% filter(Parameter1 == "Religiosity" & Parameter2 == "Social-Adaptation")
CORREL_RE_SE = CORREL %>% filter(Parameter1 == "Religiosity" & Parameter2 == "Self-Esteem")
CORREL_SA_SE = CORREL %>% filter(Parameter1 == "Social-Adaptation" & Parameter2 == "Self-Esteem")

```

`r CORREL_RE_SO$Parameter1` has a `r effectsize::interpret_r(CORREL_RE_SO$r, "cohen1988")` correlation with `r CORREL_RE_SO$Parameter2` (r = `r CORREL_RE_SO$r`, p = `r CORREL_RE_SO$p`).  
`r CORREL_RE_SE$Parameter1` has a `r effectsize::interpret_r(CORREL_RE_SE$r, "cohen1988")` correlation with `r CORREL_RE_SE$Parameter2` (r = `r CORREL_RE_SE$r`, p = `r CORREL_RE_SE$p`).  
`r CORREL_SA_SE$Parameter1` has a `r effectsize::interpret_r(CORREL_SA_SE$r, "cohen1988")` correlation with `r CORREL_SA_SE$Parameter2` (r = `r CORREL_SA_SE$r`, p = `r CORREL_SA_SE$p`).  

**[This is not autogenerated]**: Independent variables are not correlated. Regarding the dependent variable (Social-Adaptation), it has a medium correlation with self-esteem (r = .45, p < .01), and non correlation with religiosity. 

```{r}
apaTables::apa.cor.table(matched %>% select(Religiosity, `Social-Adaptation`, `Self-Esteem`), filename = here::here("outputs/tables/correlation_table.doc"), table.number = 2,show.conf.interval = F)
CORREL
# summary(CORREL, redundant = TRUE)
```

### Regressions

Since we propose a moderated mediation model, we first analyze if our main independent variable predicts variations in religiosity, the independent variable, and variations in self-esteem, the mediator. If it is not the case, then we can reject the moderated mediation model as an explanation of how these variables interact between each other.

As we can see from the outputs below, we can reject the hypothesis that Religiosity predicts Social Adaptation.

```{r}
summary(lm(`Social-Adaptation` ~ Religiosity, matched))
```

We can also reject the hypothesis that Religiosity predicts Self-Esteem.

```{r}
summary(lm(`Self-Esteem` ~ Religiosity, matched))
```

Also, group has no moderation nor mediation effect on religiosity for predicting levels of self-esteem.

```{r}
matched2 = matched %>% mutate(group = ifelse(group == "EX", 1, 0))
xz = matched$Religiosity * matched2$group
summary(lm(`Self-Esteem` ~ Religiosity + group + xz, matched2))
```

And group has no moderation nor mediation effect on religiosity for predicting levels of social adaptation.

```{r}
summary(lm(`Social-Adaptation` ~ Religiosity + group + xz, matched2))
```


As we can see, from the previous outputs, religiosity has no significant effect either on the independent variable nor on the mediator. Thus, we can say that religiosity does not predict social adaptation through self-esteem in any case.

# Exploratory Analysis

## Network correlation plot between items

The following network plot aims at discovering items between questionnaires that could be correlated between eachother. As we can see some items of the Social Adaptation Self-Evaluation Scale (SASS) do correlate with some items of Rosenberg's Self-Esteem Scale (EAR), while religiosity items do not correlate at all with either.

```{r}
tidy_cors = matched %>% select(-Estres_Percibido, -group, -Age, -sex, -edu_level) %>% 
  corrr::correlate(method = "pearson", use = "pairwise.complete.obs", quiet = TRUE) %>% 
  corrr::stretch()
```

```{r}
graph_cors <- tidy_cors %>%
  filter(abs(r) > .3) %>%
  graph_from_data_frame(directed = FALSE)

# graph_cors
```


```{r fig.height=10, fig.width=16}
plot_ggraph = 
  ggraph(graph_cors, layout = "stress") +
  geom_edge_link(aes(edge_alpha = abs(r), edge_width = abs(r), color = r)) +
  guides(edge_alpha = "none", edge_width = "none") +
  scale_edge_colour_gradientn(limits = c(-1, 1), colors = c("firebrick2", "dodgerblue2")) +
  geom_node_point(color = "#C0C0C0", size = 5, alpha = .8) +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_graph()

# Show plot
plot(plot_ggraph)

# Save plot
tiff(here::here("outputs/plots/fig2.tiff"), units="in", width=10, height=5, res=300) 
plot(plot_ggraph)
invisible(dev.off())
```


## Regression analysis between Religiosity, Self-Esteem and Perceived Stress

Since some studies that show a significant relationship between religiosity to coping and stress management, we added a perceived stress scale, and tested if religiosity predicts lower perceived stress through self-esteem.

There is no effect of religiosity over perceived stress.

```{r}
summary(lm(Estres_Percibido ~ Religiosity, matched2))
```

If we consider the experimental group, there is also no moderation nor mediation effect.

```{r}
summary(lm(Estres_Percibido ~ Religiosity + group + xz, matched2))
```

However, we can see that self-esteem does predict perceived stress

```{r}
summary(lm(Estres_Percibido ~ `Self-Esteem`, matched2))
```

From the output below we can say that self-esteem also predict social adaptation.

```{r}
xy = matched2$`Self-Esteem` * matched2$group

summary(lm(`Social-Adaptation` ~ `Self-Esteem`, matched))
```

Also, although group does not determine the strength or direction of the relationship between self-esteem and social adaptation, it does seem to play a part on explaining the process through which both variables are related. 

```{r}
summary(lm(`Social-Adaptation` ~ `Self-Esteem` + group + xy, matched2))
```

And that religiosity does predict perceived stress when we take self-esteem into account, finding both a moderation and a mediation effect.

```{r}
xw = matched2$`Self-Esteem` * matched2$Religiosity
summary(lm(Estres_Percibido ~ Religiosity + `Self-Esteem` + xw, matched2))
```

## ANOVA analysis between other Religiosity measures and Social Adaptation in non-vulnerable contexts

Since some studies that show a significant relationship between religiosity and variables related to coping and stress management use religiosity measures of religious practice and/or affiliation, we added these questions in the questionnaire that was applied to the non-vulnerable group. We now check if these variables, religious service attendance and affiliation are related to social adaptation or to self-esteem.

```{r}
DF_new_extended = data_new %>% transmute(
  id = id,
  sex = DEMOGR3_02_DIR, # 1 = Hombre; 0 = Mujer ; 2 = otro
  age = DEMOGR3_01_DIR,
  edu_level = DEMOGR3_05_DIR,
  criterio_exclusion_nac =  DEMOGR3_03_DIR, # ¿Usted nació en Chile?
  criterio_exclusion_nac10 =  DEMOGR3_04_DIR, # ¿Usted vive hace más de 10 años en Chile?
  criterio_exclusion_psi = DEMOGR3_07_DIR,
  criterio_exclusion_fis = DEMOGR3_08_DIR,
  clasificacion_gse = AIM_DIRt,
  `RS Attendance` = OTRASRELIG_11_DIR,
  Affiliation = OTRASRELIG_06_DIR,
  Autoestima = EAR_DIRt,
  SASS_Total = SASS_DIRt,
  Religiosidad = SCSORF_DIRt) %>%
  distinct(id, .keep_all = TRUE) %>%
  select(-contains("criterio"), -"clasificacion_gse", -"id") %>% mutate(group = "CN") %>% drop_na()


matched_extended = DF_new_extended %>% rename(
  "Religiosity" = "Religiosidad",
  "Self-Esteem" = "Autoestima",
  "Social-Adaptation" = "SASS_Total",
  "Age" = "age")


```

The following ANOVA analysis show that there is no difference on social adaptation nor in social adaptation levels depending on different religious service attendance levels nor type of affiliation among people who do not live in vulnerable contexts. This contradicts literature that have found that religiosity has a general positive impact on mental health variables.


#### Religious Service Attendance and Self-Esteem

```{r}
summary(aov(`Self-Esteem` ~ `RS Attendance`, data = matched_extended))
```

#### Religious Service Attendance and Social Adaptation

```{r}
summary(aov(`Social-Adaptation` ~ `RS Attendance`, data = matched_extended))
```

#### Affiliation and Self-Esteem

```{r}
summary(aov(`Self-Esteem` ~ Affiliation, data = matched_extended))
```

#### Affiliation and Social Adaptation

```{r}
summary(aov(`Social-Adaptation` ~ Affiliation, data = matched_extended))
```
